---
title: "Assingment 2 Econ320"
author: "Saier (Sam) Hu"
date: "2021/3/10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=50),tidy=TRUE)
library(tidyverse)
library(wooldridge)
library(knitr)
library(varhandle)
data('attend',ackage="wooldridge")
```

### A modification of exercise C4 chapter 3 

Use the data in ATTEND for this exercise, and and the handouts to work on these questions 

(i)	Obtain the minimum, maximum, and average values for the variables atndrte, priGPA, and ACT.  Do it this two ways to practice 

```{r, warning=FALSE}
attend %>%
  select(atndrte,priGPA,ACT)%>%
  summarise_all(list(~min(.),~max(.),~mean(.)),na.rm=TRUE)%>%
  t()%>%
  data.frame()%>%
  kable(,col.names = c("values"),digits=3)

summary(attend[,c("atndrte","priGPA","ACT")])


```


(ii)	Estimate the model $$atndrte = \beta_0 + \beta_1 priGPA + \beta_2 ACT + u$$, to make it easier create an object with the model and then show the summary. 
```{r}
mc4<-lm(formula = atndrte ~ priGPA + ACT, data = attend)

summary(mc4)

```

```{r}
mc4$coefficients

coef(mc4)
```

(iii) Calculate the predicted atndrte for all values and the mean of it. 
Then also calculate the predicted value of atndrte for the average values of the independent variables. Compare results.  
```{r}
meanyhat=mean(fitted(mc4))

means<-data.frame(priGPA = mean(attend$priGPA), 
    			ACT = mean(attend$ACT))

yhatmeans<-predict(mc4,means)

print(paste0("Average predicted attendance rate is ", round((meanyhat), 2)))
```


```{r}
print(paste0("Predicted attendance rate for average GPA and ACT is ", round((yhatmeans), 2)))

```


(iv)	What is the predicted atndrte if priGPA= 3.65 and ACT= 20? What is the predicted difference in attendant rates between this and the one for the average independent variables calculated above.  
In this case do not use the predict function for your code, retrieve the coefficients and calculate the equation $$atndrte = \hat\beta_0 + \hat\beta_1 3.65 + \hat\beta_2 20 + u$$
```{r, results='hold'}

yhat2<-coef(mc4)[1]+coef(mc4)[2]*3.65+ coef(mc4)[3]*20
names(yhat2)<-NULL # removes the names 

print(paste0("yhat= ", round((yhat2), 3)))
print(paste0("Predicted attendance rate for average GPA and ACT is ", round((yhat2-meanyhat), 3)))

```


(v) Run two more models one only for freshman another one only for sophomore. Save the results in objects mfr, msoph respectively

```{r}
mfr <- lm(atndrte ~ priGPA+ACT, subset(attend,frosh==1))
msoph <- lm(atndrte ~ priGPA+ACT,subset(attend,soph==1))


```

(vi) Using in-line code complete the table below. 

The table below compare $\beta's$, $R^2$ and observations for the models estimated. 

Variables | All | Freshmen | Sophomore
----------|-----|----------|----------
$\beta_0$ | `r round(coef(mc4)[1],3)` |  `r round(coef(mfr)[1],3)`| `r round(coef(msoph)[1],3)` 
$\beta_1$ |`r round(coef(mc4)[2],3)` |  `r round(coef(mfr)[2],3)`| `r round(coef(msoph)[2],3)` 
$\beta_2$ | `r round(coef(mc4)[3],3)` |  `r round(coef(mfr)[3],3)`| `r round(coef(msoph)[3],3)`
$R^2$ | `r round(summary(mc4)$r.squared,3)`  | `r round(summary(mfr)$r.squared,3)` |  `r round(summary(msoph)$r.squared,3)`
$N$ |`r nobs(mc4)` | `r nobs(mfr)`  | `r nobs(msoph)`


(vii) Calculate the correlation between the residuals and the priGPA for the first model. 
```{r}
round(cor(resid(mc4),mc4$model$priGPA),3)
```

## From C10 

Use the data in HTV to answer this question. The data set includes information on wages, education, parents’ education, and several other variables for 1,230 working men in 1991. 

(i)  What is the range of the educ variable in the sample? What percentage of men completed twelfth grade but no higher grade? Do the men or their parents have, on average, higher levels of education

- The range is 14 (=20-6). 41.6% of men completed twelfth grade but no higher grade. Men have higher levels of education on average than their parents do.
```{r}
# summary provides the mean and max of the variable therefore the range
summary(htv$educ)

print(paste0("Percentage of men with HS= ",round((mean(htv$educ==12)),3)))
# the mean of the logical expression gives the percentage of men that completed 12th grade but not higher. 


# See answers assignment 1 for this 
htv %>% select(educ, motheduc,fatheduc) %>% 
  summarise_all(mean) %>% 
  t()%>%
  data.frame() %>% 
  setNames(., c("Average"))%>%
  kable(digits=3)

```


(ii)  Estimate the regression model $$educ = \beta_0 + \beta_1motheduc + \beta_2fatheduc + u$$ by OLS and report the results in the usual form. How much sample variation in educ is explained by parents’ education? Interpret the coefficient on motheduc.

- 24.93% of the sample variation in educ is explained by parents’ education. Interpretation of the coefficient on motheduc: For every 1-grade increase in motheduc, there is 0.30420 grade increase in the child's education level.

(iii)  Add the variable abil (a measure of cognitive ability) to the regression from part (ii), and report the results in equation form. Does “ability” help to explain variations in education, even after controlling for parents’ education? Explain.

- Yes, “ability” does help to explain variations in education, because now 42.75% of the sample variation in educ is explained by this model.

(iv)Now estimate an equation where abil appears in quadratic form



```{r}
# model1 
model1 <- lm(formula = educ ~ motheduc + fatheduc, data = htv)
summary(model1)

# model2
model2 <- lm(formula = educ ~ motheduc + fatheduc+abil, data = htv)
summary(model2)

# model3
model3 <- lm(educ~motheduc+fatheduc+abil+I(abil^2),htv)
summary(model3)
```

(v) If you have access to a statistical program that includes graphing capabilities, use the estimates in part (iv) to graph the relationship between the predicted education and abil. Set motheduc and fatheduc at their average values in the sample, 12.18 and 12.45, respectively. See handout DGP for this. 


```{r}
dat<-data.frame(educhat=predict(model3), abil=htv$abil)
plot<-ggplot(dat , aes(x=abil,y=educhat))+ # mapping x and y 
  geom_point(col="blue", alpha=.5)+ # scatter plot defining attributes for color and transparency
 stat_function(fun=function(abil){coef(model3)[1]+ coef(model3)[2]*12.18 + coef(model3)[3]*12.45+ coef(model3)[4]*abil + coef(model3)[5]*(abil^2)}, geom="line", col=2)+ # plot the function of the relation between educ and ability
  ggtitle(paste("Predicted Education and Ability")) # title of the graph
print(plot)
```


```{r}

mean(attend$priGPA)

```

